---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

**Loading the library needed**

```{r loading, echo = T, results="hide", message=F, warning=F}
library(dplyr)
library(ggplot2)
```

**Extract and read the data**

```{r}
unzip("activity.zip")

activity <- read.csv("activity.csv", stringsAsFactors = FALSE)

activity$date <- as.Date(activity$date) #convert to 'date' format
```


## What is mean total number of steps taken per day?

**Sum of steps, grouped by date**

```{r, message=F, warning=FALSE}
activity_perday <- activity %>%
  group_by(date) %>%
  summarize(totalsteps = sum(steps))
```

**Plot a histogram of mean total number of steps taken per day**

```{r}
activity_perday %>%
  ggplot(aes(x=totalsteps)) +
  geom_histogram()
```

## What is the average daily activity pattern?

**Calculate the mean and median of steps taken per day**

```{r}
mean_median_perday <- activity %>%
  group_by(date) %>%
  summarize(totalsteps = sum(steps, na.rm = TRUE)) %>%
  summarize(mean = mean(totalsteps),
            median = median(totalsteps))

```

**Calculate the mean of steps based on certain interval**

```{r}
groupedbyinterval <- activity %>%
  group_by(interval) %>%
  summarize(mean = mean(steps, na.rm = TRUE))
```

**Plot a line plot of mean of steps based on certain interval**

```{r, message=FALSE, warning=FALSE}
groupedbyinterval %>%
  ggplot(aes(x=interval, y=mean)) + 
  geom_line()
```

**Finding the maximum interval and mean of steps**

```{r}
groupedbyinterval[which(groupedbyinterval$mean == max(groupedbyinterval$mean)),]
```


## Imputing missing values

**Calculate the sum of missing datas**

```{r}
(totalna = sum(is.na(activity$steps)))
```

**Impute the missing steps datas using the mean of the related intervals**

```{r}
activity_imp <- activity #copying the whole dataset
activity_imp$steps_imp <- activity_imp$steps #copy original steps to new column

for (x in 1:length(activity_imp$steps_imp)){
    if(is.na(activity_imp$steps[x]) == TRUE){
      activity_imp$steps_imp[x] <- round(groupedbyinterval$mean[which(groupedbyinterval$interval == activity_imp$interval[x])])
    }
}
```

**Calculate the sum of steps per day using imputed values from code chunk before**

```{r}
activity_imp_perday <- activity_imp %>%
  group_by(date) %>%
  summarize(totalsteps = sum(steps_imp))
```

**Plot a histogram of mean total number of steps taken per day using the imputed datas**

```{r}
activity_imp_perday %>%
  ggplot(aes(x=totalsteps)) +
  geom_histogram()
```

****Calculate the mean and median of steps taken per day using the imputed datas**

```{r}
mean_median_imp_perday <- activity_imp %>%
  group_by(date) %>%
  summarize(totalsteps = sum(steps_imp, na.rm = TRUE)) %>%
  summarize(mean = mean(totalsteps),
            median = median(totalsteps))
```


## Are there differences in activity patterns between weekdays and weekends?

**Calculate the mean of steps taken, grouped by Weekend or Weekday**

```{r}
activity_imp$dayend <- ifelse(weekdays(activity_imp$date) == "Saturday" | weekdays(activity_imp$date) == "Sunday", 
                              "Weekend", 
                              "Weekday")

activity_imp_grouped_dayend <- activity_imp %>%
  group_by(dayend, interval) %>%
  summarize(mean = mean(steps_imp, na.rm = TRUE))

```

**Plot a line plot of mean of steps based on certain interval using the imputed datas, grouped by Weekend or Weekday**

```{r}
activity_imp_grouped_dayend %>%
  ggplot(aes(x=interval, y=mean)) +
  geom_line() +
  facet_wrap(~dayend)
```